Tracking and predicting turn structure during language acquisition

### 2014-09-29 paper/CasillasFrank-Anticip.tex
Merging in MCF's edits

Conversation distinguishes itself from non-conversational turn-taking behaviors by the complexity of the turn sequencing involved. In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk. For example, conversational turns come grouped into semantically-contingent sequences of action. The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!''). Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}). To take a turn, participants need to make predictions about what types of talk will be relevant next. In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from, or no obvious next action at all (e.g., after a closed sequence).

Despite this complexity, conversational turn-taking is precise in its timing. Across a diverse sample of conversations in 10 languages, one study found a consistent average transition time of 0--200 msec at points of speaker switch \citep{stivers2009}. Current models of speech production suggest that it takes approximately 600 msec to produce a word, and even longer to produce a simple utterance \citep{levelt1989}. So in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended. To formulate their response early on, speakers must track and anticipate what types of response might become relevant next. They also need to predict the content and form of upcoming speech so that they can launch their articulation at exactly the right moment. Prediction thus plays a key role in timely turn-taking.

\subsection{Adults' turn prediction}

Adults have a lot of information that could potentially help with their predictions about upcoming turn structure. Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreiraUndRev}. Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}. The sequential context of a turn can make it clear what will come next: answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}. Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure, while prosodic information plays more of a secondary role \citep{de-ruiter2006}. For example, a mid-high ``continuation'' tone can cancel the possibility of a speaker switch at an otherwise syntactically- and pragmatically-complete turn point (e.g., at the start of a story-telling; \citealp{ford1996, torreiraUndRev}).

@@

In another study, \citet*{torreiraUndRev} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-boundary identification. They found that participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete unit without a boundary tone. Utterance-level prosodic boundaries embedded in multi-utterance turns \textit{could} trick participants into anticipating a turn end 29\% of the time, however (see also de \citet{de-ruiter2006}:525). Their findings corroborate other corpus and experimental findings promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.

In sum, adults accurately and spontaneously make predictions about upcoming turn structure. They use a wide variety of cues in making their predictions and then act immediately upon them. Their predictive processing relies on a sophisticated body of knowledge about linguistic structure, non-verbal signals, and social actions. The acquisition of turn-taking skills must then be tied to children's knowledge about language, gaze, gesture, and social cues. But children's turn taking starts early in infancy, long before first words or gestures emerge; a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn-taking.

@@

The timing of children's responses to their caregivers' speech shows a non-linear pattern of fall-rise-fall from early infancy to middle childhood. Infants' turn timing at three months is often too early or too late: they start vocalizing in overlap on nearly 40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average latency of 675 msec (adult average: 200 msec). Between four and nine months, children begin to reduce the number of turns happening overlap while also improving on their average response latency. But then, later on, children's response latencies slow down again, peaking at average gaps of 1100 msec at nine months, with only very gradual improvement after that \citep{hilbrinkInPrep}. While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next several years \citep{ervin-tripp1979}.

@@

Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds, finding that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults. Their study was the first to focus on the role of linguistic processing in children's turn predictions. They showed their participants two types of conversation videos: one normal and one with flattened pitch (i.e., with removed intonation contours), finding that only 36-month-olds were affected by a lack of intonation. The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}. Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 3;0), especially with respect to semantic access. They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 3;0).

@@

Second, the conversation stimuli they used were somewhat unusual. The average gap between turns was 900 msec, which is much longer than typical adult timing, where gaps average around 200 msec \citep{stivers2009}. The speakers in the videos were also asked to minimize their movements while performing a scripted, adult-directed conversation, which would have created a somewhat unnatural stimulus. In addition, in order to produce more naturalistic conversation, it would have been ideal to localize sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers). But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the speakers. Again, we attempt to address these issues in our current study.

@@

\subsection{The Current Study}

We report here on the role of linguistic processing in children's predictions about upcoming turn structure. In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling for the amount of lexicosyntactic and prosodic information available. Experiment 1 uses a paradigm in which English-speaking participants view video clips of naturalistic conversations from several different languages, to create conditions under which lexicosyntactic information was differentially available, but prosodic and non-linguistic cues were intact. This paradigm showed minimal differences between the predictive looking behavior of preschoolers and adults. In Experiment 2, we created artificial (puppet) visual stimuli to better separate these linguistic cues from one another. In this more controlled paradigm, we found that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.

@@

We recorded participants' eye movements as they watched six short videos of two-person (dyadic) conversation. Each video featured an improvised discourse in one of five languages (English, German, Hebrew, Japanese, and Korean). The participants, all native English speakers, were only expected to understand the two videos in English. We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., prosody, gaze, breath, phrase final lengthening). Using this method, we compared children and adult's anticipatory looks from the current speaker to the upcoming speaker at points of turn transition in English and non-English videos.

@@

\textit{Video recordings}. We recorded pairs of talkers while they conversed in a sound-attenuated booth (see sample frame in Figure \ref{fig:speakers}). Each talker was a native speaker of the language being recorded, and each talker pair was male-female. Using a Marantz PMD 660 solid state field recorder, we captured audio from two lapel microphones, one attached to each participant, while simultaneously recording video from the built-in camera of a MacBook. The talkers were volunteers, all members of the local post-graduate community, and were acquainted with their recording partner ahead of time.

@@

After recording, we hand-aligned the audio and video files, and cropped each recording to the 30-second interval with the most turn activity. Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback. This gave each voice in the videos a localized source (from the left or right loudspeaker). We coded each turn transition in the videos for language group (English vs. non-English), inter-turn gap duration (in milliseconds), and speech act (question vs. non-question). The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which don't. The coding for speech act in non-English videos was completed independently by two coders and then checked and discussed for agreement.

Because the conversational stimuli were recorded semi-spontaneously, the duration turn transitions and the number of speaker transitions in each video was variable. We measured the duration of each turn transition from the audio recording associated with each video. We excluded turn transitions longer than 550 msec and shorter than 0 msec (those with transitional overlap) from analysis.\footnote{Overlap occurs when an addressee begins a new turn before the current turn is finished. When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response begins \textit{earlier} than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}. As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$22) and non-English (N$=$21) videos. On average, the inter-turn gaps for English videos (mean$=$273 msec, median$=$291 msec) were slightly longer than for non-English videos (mean$=$226 msec, median$=$205 msec). The longer gaps in the English videos could give them a slight advantage: our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.

@@

Using these criteria, we created a three-step algorithm for detecting anticipatory gaze switches. Consider, for example, when a child observes a turn transition between speaker A and speaker B (Figure \ref{fig:criterion}). First, the child would need to fixate on A for at least 100 msec at the end of A's turn. But what is the relevant window for fixation? Consistent with the distribution of responses in explicit turn-boundary prediction tasks (de Ruiter et al., 2006), we counted any early switches as ``truly turn-related'' as long as they occurred in the final 333 msec of A's turn (200 msec for adults). To allow for these early switches, the window for fixating on A must start 433 msec before the end of A's turn (300 msec for adults). Second, given that the child fixated on A, we checked to see if the child also fixated on B. Because we allowed for early switches, the first possible fixation on B could start 333 msec before the end of A's turn. Assuming that it took the child 333 msec to plan an eye movement, the latest point when the gaze could have \textit{anticipated} speech would be 333 msec after B's turn began---a switch after this point could simply be a reaction to the onset of B's speech. Thus, the relevant window for fixating on B started 333 msec before the gap and ended 433 msec after it (200 msec and 300 ms, respectively, for adults). Success on the first two criteria jointly implied that the child made an anticipatory switch in gaze. Our third step was to estimate when the switch occurred. Starting at a point when we knew participants were fixated on A (from step 1), we incremented forward in time to the first gaze measurement where the child looked away from A. We recorded the look-away timestamp as the time of gaze launch to the upcoming speaker.

\paragraph{Random baseline analysis} To ensure that participants' looking behavior was turn-related and not simply the result of random switching, we corrected for the probability of making an anticipatory switch at random. We estimated the baseline chance of making an anticipatory switch at each turn transition for each person. We ran all participants' eye-tracking data through switch identification with 100 randomly-shuffled versions of the original turn transition windows (Figure \ref{fig:shuffling}). For each of the 100 randomly-shuffled versions, we recorded and then averaged whether or not each participant made an anticipatory switch at each ``turn transition.'' We then subtracted the random baseline value from each participant's actual switch result (1 or 0) for each turn transition to arrive at the final, baseline-corrected values. Subtracting the random baseline this way can give us an estimate of how often participants switched above chance across ages, conditions, and transition types. If participants made the same number of anticipatory switches, regardless of turn transition location, there would be no evidence that their switching behavior was linked to the conversation stimulus.

@@

We identified anticipatory gaze switches for all 43 usable turn transitions, based on the criteria and algorithm outlined in Section \ref{sec:algorithm}. Because participants had greater access to lexical and syntactic information in English, we expected to see greater anticipation in the English videos than in the non-English videos. We also predicted that anticipation would be greater following questions compared to non-questions; questions have early cues to upcoming turn transition (e.g., \textit{wh-} words, subject-auxiliary inversion), and also make a next response immediately relevant. We also predicted that anticipatory looks would increase with development, along with children's increased linguistic competence.

To estimate the effects of linguistic condition and age on switching behavior, we corrected each participant's data by subtracting their random baseline estimations from their actual switching data for each turn transition. All comparisons below are made with the baseline-corrected values.

@@

Participants at all ages made anticipatory gaze switches more often than would be expected by chance, whether they could understand the language in the video or not (Figure \ref{fig:randvsrealFL}). Anticipatory switch behavior in the non-English videos could not have been based on lexicosyntactic information, so both adults and children must have relied on other sources of information to make their predictions (e.g., prosody and non-verbal signals).

@@

Participants also showed an advantage for questions. Children made anticipatory switches twice as often when they heard a question, compared to a non-question (18.7\% and 7.3\%). But the advantage for questions was not universal---it depended on language condition and age. The advantage for questions didn't emerge until age five in the non-English videos (Figure \ref{fig:questionsFL}).

@@

We quantified these findings using two linear mixed-effects models of baseline-corrected anticipatory switches \citep{lme4, R}. We built one model each for children and adults. The child model included condition (English vs. non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot compute reliable cross-language differences. Gaze behavior might be best for languages that have more structural overlap English. English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \cite{carlson2005}. We would need multiple items from each of the language to check for similarity effects of specific linguistic features.}, transition type (question vs. non-question), age, and duration of the inter-turn gap as predictors, with full interactions between condition, transition type, and age.  We included the duration of the inter-turn gap as a predictor since longer gaps also provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}). We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}. The adult model was exactly the same as the children's, excluding age-related effects.

Children's anticipatory gaze switches showed a main effect of language condition (\textit{$\beta$}=-0.448, \textit{SE}=0.126, \textit{t}=-3.56, \textit{p}$<$.001).\footnote{We derive \textit{p}-values here by treating \textit{t} and \textit{z} (with a two-tailed assumption), as is justified for large data sets \citep{barr2013}.} As predicted, children made more anticipatory looks while they were watching videos that they could understand. In the English videos, children had access to lexicosyntactic information, which could have helped them make better predictions about upcoming turn structure. Although children performed better on the English videos, lexicosyntactic information was not exclusively responsible for their anticipatory switches; children still performed remarkably well when no lexical information was present at all, and even three-year-olds made anticipatory switches at an above-chance rate in the non-English videos.

@@

Finally, we had predicted that children would show early advantages for prosodic cues over syntactic ones because children begin developing language-specific knowledge about prosody long before lexicosyntax. We weren't able to directly test this prediction in Experiment 1 because we used a limited range of ages (three- to five-year-olds) and because we only controlled for lexicosyntactic information. Children had prosodic cues in both English and non-English videos, although the prosodic signal was non-native in the latter case. So, although using non-English videos was a natural way to control for lexical information, it did not allow us to assess the role of prosody. Relatedly, the advantage for questions was greater in the English videos than the non-English videos, suggesting that lexicosyntactic cues are important for children in identifying questions during conversation. Again, to test this idea, we would need to directly compare lexicosyntactic and prosodic cues in the participants' native language. In sum, the current study lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.

@@

We improved our design by using native-language stimuli, controlling for lexical \textit{and} prosodic information, eliminating non-verbal cues, and testing children from a wider age range. All of the videos in Experiment 2 were in the participants' native language (American English). To tease apart the role of lexical and prosodic information, we phonetically manipulated the speech signal for pitch, rhythm, and lexical access. By testing one- to six-year-olds we hoped to find the developmental onset of turn-predictive gaze. We also hoped to measure changes in the relative roles of prosody and lexicosyntax across development.

Non-verbal cues in Experiment 1 (e.g., gaze and gesture) could have helped participants make predictions about upcoming turn structure  \citep{rossano2009, stivers2010}. Since our focus is on linguistic cues, we eliminated all gaze and gestural signals in Experiment 2 by replacing the videos of human actors with videos of puppets. Puppets are less realistic and expressive than human actors but, since we wanted to eliminate gestural and gaze cues, the puppets created a natural context for having somewhat motionless talkers in the videos. Additionally, the prosody-controlled condition included small but global changes to syllable duration that would have required complex video manipulation or precise re-enactment with human talkers, neither of which was feasible. For these reasons, we decided to substitute puppet videos for human videos in the final stimuli.

@@

\textit{Video recordings}. We created puppet video recordings to match the manipulated 20-second audio clips. The puppets were minimally expressive; the experimenter could only control the opening and closing of their mouths; their head, eyes, arms, and body stayed still. Puppets were positioned looking forward to eliminate shared gaze as a cue to turn structure \citep{thorgrimssonUndRev}. We took care to match the puppets' mouth movements to the syllable onsets as closely as possible and avoided any mouth movement before the onset of a turn. We then hand-aligned the manipulated audio clips to the puppet video recordings.

@@

Participants at all ages made anticipatory gaze switches more often than would be expected by chance in almost all conditions, but sometimes by a small margin (Figure \ref{fig:randvsrealEN}). For example, anticipatory switches in the \textit{no speech} videos were often only modestly above chance (with the exception of 5- and 6-year-olds). Notably, while children performed well with \textit{prosody only} speech from the start, they seemed to only reliably exceed chance in the \textit{words only} condition at ages 3;0 and older, giving some support to the idea that there is an early advantage for prosody over lexicosyntax in children's turn prediction.

>>>>

Conversation distinguishes itself from non-conversational turn-taking behaviors by the complexity of the turn sequencing involved. In the examples above (traffic, games, and monkeys) the set of sequence and action types is far more limited and predictable than what we find in everyday talk. For example, conversational turns come grouped into semantically-contingent sequences of action. The groups can span turn-by-turn exchanges (e.g., simple question--response, ``How are you?''--``Fine.'') or sequence-by-sequence exchanges (e.g., reciprocals, ``How are you?''--``Fine, and you?''--``Great!''). Sequences of action drive the conversation forward into the next, relevant sequences of talk (e.g., ''And you?''--``Great!''--``Why's that?''; \citealp{schegloff2007}). To take a turn, participants need to make predictions about what conversational content will be relevant next. In some cases, relevant next turns are somewhat obvious (e.g., question--response) while, in other cases, there are multiple relevant next actions to choose from, or no obvious next action at all (e.g., after a closed sequence).

Despite this complexity, conversational turn-taking is often precise in its timing. Across a diverse sample of conversations in 10 languages, one study found a consistent average transition time of 0--200 msec at points of speaker switch \citep{stivers2009}. Experimental results and current models of speech production suggest that it takes approximately 600 msec to produce a content word, and even longer to produce a simple utterance \citep{griffin2000, levelt1989}. So in order to achieve 200 msec turn transitions, speakers must begin formulating their response before the prior turn has ended. Moreover, to formulate their response early on, speakers must track and anticipate what types of response might become relevant next. They also need to predict the content and form of upcoming speech so that they can launch their articulation at exactly the right moment. Prediction must thus play a key role in timely turn-taking.

\subsection{Adults' turn prediction}
% FIX: compress deR: make it clear about prosody issues, perhaps talk more clearly about torreira
Adults have a lot of information at their disposal that could potentially help with their predictions about upcoming turn structure. Lexical, syntactic, and prosodic information (e.g., \textit{wh}- words, subject-auxiliary inversion, list intonation) can all inform addressees about upcoming linguistic structure \citep{de-ruiter2006, duncan1972, ford1996, torreiraUndRev}. Non-verbal cues (e.g., gaze, posture, and pointing) often appear at turn-boundaries as late indicators of an upcoming speaker switch \citep{rossano2009, stivers2010}. The sequential context of a turn can make it clear what will come next: answers after questions; thanks or denial after compliments, et cetera \citep{schegloff2007}. Prior work suggests that adult listeners primarily use lexicosyntactic information to accurately predict upcoming turn structure, while prosodic information plays more of a secondary role \citep{de-ruiter2006}. For example, a mid-high ``continuation'' tone can cancel the possibility of a speaker switch at an otherwise syntactically- and pragmatically-complete turn point (e.g., at the start of a story-telling; \citealp{ford1996, torreiraUndRev}).

@@

In another study, \citet*{torreiraUndRev} used a combination of button-press and verbal responses to investigate the relationship between lexicosyntactic and prosodic cues in turn-boundary identification. They found that participants never verbally responded or pressed the ``turn-end'' button when hearing a syntactically-complete unit without an intonational phrase boundary. However, utterance-level prosodic boundaries embedded in multi-utterance turns could trick participants into anticipating a turn end 29\% of the time (see also de \citet{de-ruiter2006}:525). Their findings corroborate other corpus and experimental findings promoting a combination of cues (lexicosyntactic, prosodic, and pragmatic) as key for accurate turn-end prediction \citep{duncan1972, ford1996, hirvenkari2013}.

In sum, adults accurately and spontaneously make predictions about upcoming turn structure. They use a wide variety of cues in making their predictions and then act immediately upon them. Their predictive processing relies on a sophisticated body of knowledge about linguistic structure, non-verbal signals, and social actions. The later acquisition of verbal turn-taking skills must then be tied to children's knowledge about language, gaze, gesture, and social cues. But children's turn taking starts early in infancy, long before first words or gestures emerge; a primary role for lexicosyntactic cues doesn't fit well with children's pre-verbal turn-taking.

@@

The timing of children's responses to their caregivers' speech shows a non-linear pattern of fall-rise-fall from early infancy to middle childhood. A recent study by Hilbrink et al. (in preparation) finds that infants' turn timing at three months is often too early or too late: they start vocalizing in overlap on nearly 40\% of their caregivers' turns, and their non-overlapped vocalizations come after an average latency of 675 msec (adult average: 200 msec). Between four and nine months, children begin to reduce the number of turns happening overlap while also improving on their average response latency. But then, later on, children's response latencies slow down again, peaking at average gaps of 1100 msec at nine months, with only very gradual improvement after that \citep{hilbrinkInPrep}. While children's avoidance of overlap is nearly adult-like by nine months, the timing of their non-overlapped responses stays much longer than the 200 msec standard for the next several years \citep{ervin-tripp1979}.

@@

Keitel and colleagues \citeyearpar{keitel2013} addressed the random baseline issue in their study of 6-, 12-, 24-, and 36-month-olds. They asked participants to watch short videos of conversation, and tracked their eye movements at points of speaker change. They found that children's anticipatory gaze frequency was only greater than chance for 36-month-olds and adults. Their study was the first to focus on the role of linguistic processing in children's turn predictions. They showed their participants two types of conversation videos: one normal and one with flattened pitch (i.e., with flattened intonation contours), finding that only 36-month-olds were affected by a lack of intonation contours. The adult control group made equal numbers of anticipatory looks in the videos, with and without intonation contours, consistent with prior adult findings \citep{de-ruiter2006}. Keitel and colleagues concluded that children's ability to predict upcoming turn structure relies on their ability to comprehend the stimuli (emerging around 3;0), especially with respect to semantic access. They also suggest that intonation takes a secondary role in turn prediction, but only \textit{after} children acquire more sophisticated, adult-like language comprehension systems (sometime after 3;0).

@@

Second, the conversation stimuli they used were somewhat unusual. The average gap between turns was 900 msec, which is much longer than typical adult timing, where gaps average around 200 msec \citep{stivers2009}. The speakers in the videos were also asked to minimize their movements while performing a scripted and adult-directed conversation, which would have created a somewhat unnatural stimulus. In addition, in order to produce more naturalistic conversation, it would have been ideal to localize sources for the two voices in the video (i.e., to have the voices come out of separate left and right speakers). But both voices were recorded and played back on the same audio channel, which may have made it more difficult to distinguish the speakers. Again, we attempt to address these issues in our current study.

@@

\subsection{The Current Study}
% add bit about children's prosodic development and link this to the plans for the experiments
% mention that we will be looking at anticipatory gaze switching
We report here on the role of linguistic processing in children's predictions about upcoming turn structure. In two eye-tracking experiments, we measured children's anticipatory gaze to upcoming responders while controlling for the amount of lexicosyntactic and prosodic information available. Experiment 1 uses a paradigm in which English-speaking participants view video clips of naturalistic conversations from several different languages, to create conditions under which lexicosyntactic information was differentially available, but prosodic and non-linguistic cues were intact. This paradigm showed minimal differences between the predictive looking behavior of preschoolers and adults. In Experiment 2, we created artificial (puppet) visual stimuli to better separate these linguistic cues from one another. In this more controlled paradigm, we found that children's predictive looking behavior improved from ages one to six, but that even one-year-olds made more anticipatory looks than would be expected by chance.

@@

We recorded participants' eye movements as they watched six short videos of two-person (dyadic) conversation. Each video featured an improvised discourse in one of five languages (English, German, Hebrew, Japanese, and Korean). The participants, all native English speakers, were only expected to understand the two videos in English. We showed participants non-English videos to limit their access to lexical information while maintaining their access to other cues to turn boundaries (e.g., (non-native) prosody, gaze, breath, phrase final lengthening). Using this method, we compared children and adult's anticipatory looks from the current speaker to the upcoming speaker at points of turn transition in English and non-English videos.

@@

\textit{Video recordings}. We recorded pairs of talkers while they conversed in a sound-attenuated booth (see sample frame in Figure \ref{fig:speakers}). Each talker was a native speaker of the language being recorded, and each talker pair was male-female. Using a Marantz PMD 660 solid state field recorder, we captured audio from two lapel microphones, one attached to each participant, while simultaneously recording video from the built-in camera of a MacBook laptop computer. The talkers were volunteers, all members of the local post-graduate community, and were acquainted with their recording partner ahead of time.

@@

After recording, we hand-aligned the audio and video files, and cropped each recording to the 30-second interval with the most turn activity. Because we recorded the conversations in stereo, the male and female voices came out of separate speakers during video playback. This gave each voice in the videos a localized source (from the left or right loudspeaker). We coded each turn transition in the videos for language group (English vs. non-English), inter-turn gap duration (in milliseconds), and speech act (question vs. non-question). The non-English stimuli were coded for speech act from a monolingual English-speaker's perspective, i.e., which turns ``sound like'' questions, and which don't.

Because the conversational stimuli were recorded semi-spontaneously, the duration of turn transitions and the number of speaker transitions in each video was variable. We measured the duration of each turn transition from the audio recording associated with each video. We excluded turn transitions longer than 550 msec and shorter than 0 msec (those with transitional overlap) from analysis.\footnote{Overlap occurs when an addressee begins a new turn before the current turn is finished. When overlap occurs, observers cannot switch their gaze in anticipation of the response because the response begins \textit{earlier} than expected; participants expect conversations to proceed with ``one speaker at a time'' \citep{sacks1974}. As such, they would still be fixated on the prior speaker when the overlap started, and then would have to switch their gaze \textit{reactively} to the responder.} This left approximately equal numbers of turn transitions available for analysis in the English (N$=$22) and non-English (N$=$21) videos. On average, the inter-turn gaps for English videos (mean$=$273 msec, median$=$291 msec) were slightly longer than for non-English videos (mean$=$226 msec, median$=$205 msec). The longer gaps in the English videos could give them a slight advantage: our definition of an ``anticipatory gaze shift'' includes shifts that are initiated during the gap between turns (Figure \ref{fig:criterion}), so participants had slightly more time to make anticipatory shifts in the English videos.

@@

% check over again to make as clear as possible
Using these criteria, we created a three-step algorithm for detecting anticipatory gaze switches. Consider, for example, when a child observes a turn transition between speaker A and speaker B (Figure \ref{fig:criterion}). First, the child would need to fixate on A for at least 100 msec at the end of A's turn. But what is the relevant window for fixation? Consistent with the distribution of responses in explicit turn-boundary prediction tasks (de Ruiter et al., 2006), we counted any early switches as ``truly turn-related'' as long as they occurred in the final 333 msec of A's turn (200 msec for adults). To allow for these early switches, the window for fixating on A must start 433 msec before the end of A's turn (300 msec for adults). Second, given that the child fixated on A, we checked to see if the child also fixated on B. Because we allowed for early switches, the first possible fixation on B could start 333 msec before the end of A's turn. Assuming that it took the child 333 msec to plan an eye movement, the latest point when the gaze could have \textit{anticipated} speech would be 333 msec after B's turn began---a switch after this point could simply be a reaction to the onset of B's speech. Thus, the relevant window for fixating on B started 333 msec before the gap and ended 433 msec after it (200 msec and 300 ms, respectively, for adults). Success on the first two criteria jointly implied that the child made an anticipatory switch in gaze. Our third step was to estimate when the switch occurred. Starting at a point when we knew participants were fixated on A (from step 1), we incremented forward in time to the first gaze measurement where the child looked away from A. We recorded the look-away timestamp as the time of gaze launch to the upcoming speaker.

% check over again to make as clear as possible
\paragraph{Random baseline analysis} To ensure that participants' looking behavior was turn-related and not simply the result of random switching, we corrected for the probability of making an anticipatory switch at random. We estimated the baseline chance of making an anticipatory switch at each turn transition for each person. We ran all participants' eye-tracking data through switch identification with 100 randomly-shuffled versions of the original turn transition windows (Figure \ref{fig:shuffling}). For each of the 100 randomly-shuffled versions, we recorded and then averaged whether or not each participant made an anticipatory switch at each ``turn transition.'' We then subtracted the random baseline value from each participant's actual switch result (1 or 0) for each turn transition to arrive at the final, baseline-corrected values. Subtracting the random baseline this way can give us an estimate of how often participants switched above chance across ages, conditions, and transition types. If participants made the same number of anticipatory switches, regardless of turn transition location, there would be no evidence that their switching behavior was linked to the conversation stimulus.

@@

We identified anticipatory gaze switches for all 43 usable turn transitions, based on the criteria and algorithm outlined in Section \ref{sec:algorithm}. Because participants had greater access to linguistic information in English, we expected to see greater anticipation in the English videos than in the non-English videos. We also predicted that anticipation would be greater following questions compared to non-questions; questions have early cues to upcoming turn transition (e.g., \textit{wh-} words, subject-auxiliary inversion), and also make a next response immediately relevant. We also predicted that anticipatory looks would increase with development, along with children's increased linguistic competence.

@@

To estimate the effects of linguistic condition and age on switching behavior, we estimated each participants' random baseline switching behavior for each turn transition. Participants at all ages made anticipatory gaze switches more often than would be expected by chance, whether they could understand the language in the video or not (Figure \ref{fig:randvsrealFL}). Anticipatory switch behavior in the non-English videos could not have been based on lexicosyntactic information, so both adults and children must have relied on other sources of information to make their predictions (e.g., non-native prosodic cues and non-verbal signals). We then corrected each participant's data by subtracting their random baseline estimations from their actual switching data for each turn transition. All comparisons below are made with the baseline-corrected values.

@@

Participants also showed an advantage for questions (Figure \ref{fig:questionsFL}). Children made anticipatory switches twice as often when they heard a question, compared to a non-question (18.7\% and 7.3\%). But the advantage for questions was not universal---it depended on language condition and age. The advantage for questions didn't emerge until age five in the non-English videos .

@@

We quantified these findings using two linear mixed-effects models of participants' baseline-corrected anticipatory switches \citep{lme4, R}. We built one model each for children and adults. The child model included condition (English vs. non-English)\footnote{Because each non-English language was represented by a single stimulus, we cannot compute reliable cross-language differences. Gaze behavior might be best for languages that have more structural overlap English. English speakers can make predictions about the strength of upcoming Swedish prosodic boundaries nearly as well as Swedish speakers do, but Chinese speakers are at a disadvantage in the same task \cite{carlson2005}. We would need multiple items from each of the language to check for similarity effects of specific linguistic features.}, transition type (question vs. non-question), age, and duration of the inter-turn gap as predictors, with full interactions between condition, transition type, and age.  We included the duration of the inter-turn gap as a predictor since longer gaps also provide more opportunities to make anticipatory switches (Figure \ref{fig:criterion}). We additionally included random effects of item (turn transition) and participant, with random slopes of condition, transition type, and their interaction for participants \citep{barr2013}. The adult model was exactly the same as the children's, excluding age-related effects.

Children's anticipatory gaze switches showed a main effect of language condition (\textit{$\beta$}=-0.448, \textit{SE}=0.126, \textit{t}=-3.56, \textit{p}$<$.001).\footnote{We derive \textit{p}-values here by treating \textit{t} as \textit{z} (with a two-tailed assumption), as is justified for large data sets \citep{barr2013}.} As predicted, children made more anticipatory looks while they were watching videos that they could understand. In the English videos, children had access to lexicosyntactic information, which could have helped them make better predictions about upcoming turn structure. Although children performed better on the English videos, lexicosyntactic information was not exclusively responsible for their anticipatory switches; children still performed remarkably well when no lexical information was present at all, and even three-year-olds made anticipatory switches at an above-chance rate in the non-English videos.

@@

Finally, we had predicted that children would show early advantages for prosodic cues over syntactic ones because children begin developing language-specific knowledge about prosody long before lexicosyntax. We weren't able to directly test this prediction in Experiment 1 because we used a limited range of ages (three- to five-year-olds) and because we only controlled for lexicosyntactic information. Children had prosodic cues in both English and non-English videos, although the prosodic signal was non-native in the latter case. So, although using non-English videos was a natural way to control for lexical information, it did not allow us to assess the role of prosody. Relatedly, the advantage for questions was greater in the English videos than the non-English videos, suggesting that lexicosyntactic cues are important for children in identifying questions during conversation. Again, to test this idea, we would need to directly compare lexicosyntactic and prosodic cues in the participants' native language. In sum, Experiment 1 lays the analytic groundwork for a method that allows for greater experimental control, which we introduce in Experiment 2.

@@

We improved our design by using native-language stimuli, controlling for lexical \textit{and} prosodic information, eliminating non-verbal cues, and testing children from a wider age range. All of the videos in Experiment 2 were in the participants' native language (American English). To tease apart the role of lexical and prosodic information, we phonetically manipulated the speech signal for pitch, syllable duration, and lexical access. By testing one- to six-year-olds we hoped to find the developmental onset of turn-predictive gaze. We also hoped to measure changes in the relative roles of prosody and lexicosyntax across development.

Non-verbal cues in Experiment 1 (e.g., gaze and gesture) could have helped participants make predictions about upcoming turn structure  \citep{rossano2009, stivers2010}. Since our focus is on linguistic cues, we eliminated all gaze and gestural signals in Experiment 2 by replacing the videos of human actors with videos of puppets. Puppets are less realistic and expressive than human actors, but they create a natural context for having somewhat motionless talkers in the videos (thereby allowing us to eliminate gestural and gaze cues). Additionally, the prosody-controlled condition included small but global changes to syllable duration that would have required complex video manipulation or precise re-enactment with human talkers, neither of which was feasible. For these reasons, we decided to substitute puppet videos for human videos in the final stimuli.

@@

\textit{Video recordings}. We created puppet video recordings to match the manipulated 20-second audio clips. The puppets were minimally expressive; the experimenter could only control the opening and closing of their mouths; their head, eyes, arms, and body stayed still. Puppets were positioned looking forward to eliminate shared gaze as a cue to turn structure \citep{thorgrimssonUndRev}. We took care to match the puppets' mouth movements to the syllable onsets as closely as possible and avoided any mouth movement before the onset of a turn. We then added the manipulated audio clips to the puppet video recordings by hand.

@@

% address this!!
Participants at all ages made anticipatory gaze switches more often than would be expected by chance in almost all conditions, but sometimes by a small margin (Figure \ref{fig:randvsrealEN}). For example, anticipatory switches in the \textit{no speech} videos were often only modestly above chance (with the exception of 5- and 6-year-olds). Notably, while children performed well with \textit{prosody only} speech from the start, they seemed to only reliably exceed chance in the \textit{words only} condition at ages 3;0 and older, giving some support to the idea that there is an early advantage for prosody over lexicosyntax in children's turn prediction.
